### 2024-09-11 texts/paper_closing.tex
polish conclusion, minor section moves, add US nbrs

% Created 2024-09-10 di 19:08
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

@@

Neither gender of the founder nor external shocks (the Great Recession and the Covid-19 Pandemic) are associated with significantly higher closing rates.
\paragraph*{Founder dependence (H2)}

@@

@@

While so far little evidence points toward low diversification, i.e. reliance on the founder, as a key vulnerability of private museums to unvoluntary closures, the centrality of the founder may be more pronounced in voluntary closures. 
Indicated by unconventional naming choice, a subset of founders might from the very beginning envision the museum as temporary.
On the other hand, higher closing rates for unconventionally named museums might reflect absence of recognition and support by third parties in the environment.
While the question of which mechanism explains the association between name and survival, the role of the environment can be clearly seen in terms of nearby institution counts and population measures, with both factors leading to higher closing rates.

While overall private museum closing rates are generally not diverging from museums in general, and may even lie below the average, it has to be kept in mind that this comparison is based on American non-profit museums.
To the best of my knowledge, currently no data sources exist that allow estimation of closing rates of (European) public, state-funded museums.
While public museums have been characterized to some extent as very stable due to government picking up deficits \parencite{Meier_Frey_2003_faces,Bechtler_Imhof_2018_future}, \textcite{Walker_2019_collector} argues that due to declining cultural spending even public museums might face hard times.
Most national public flagship institutions will presumably remain sufficiently funded to survive, however this might not be the case for the much larger of smaller museums in less touristic regions who rely on more limited municipal funding (49\% of the European museum population, \cite{EGMUS_2024_complete}), as \textcite{StylianouLambert_etal_2014_museums} explore for a number of smaller publicly funded ethnography museums in Cyprus.
Yet without access to museum registers comparable to the IRS/NCCS datasets, questions regarding the relation between type of ownership and survival remain speculative.

\subsection*{Conclusion}

Private museum have been characterized as fragile institutions due to their reliance on a single founder.
However, my analysis finds neither a higher substantially higher mortality rates for private museums compared to (art) museums generally, nor a conclusive relationship between museum closure and a number of founder characteristics, such as founder gender, death, exclusion from the Artnews collector ranking or inclusion of founder name in the museum name.
While Self-identification and audience/competition are significantly associated with higher mortality, and possibly reflect or influence founder motivations, they are also subject to alternative mechanisms, namely devaluation by third parties and competition, respectively (self-identification furthermore concerns a relatively small group of institutions).
Furthermore, even if they reflect founder motivations, these are likely cases of voluntary closure; there is thus no large evidence that private museums are at a higher risk of unvoluntary closure than more diversified museums as a consequence of their reliance on a single founder. 
\subsubsection*{Limitations}

For this study, I only investigated the occurence of closure events.
However, private museums can stop existing as private museums not just by closing down, but also by being transformed into other organizational forms.
One way in which this can happen is when other parties, such as local governments other philanthropists, foundations, or corporations become substantially involved in the ownership, governance and/or funding of the museum.
Sometimes this involves changes in the composition of the board of directors (or equivalent), with new parties gaining seats and hence influence in decision-making, thereby diminishing the private character of the museum to an extent that it no longer fits the working definition.
Museums which experienced this trajectory have been excluded from the analysis as explorative investigation showed that it was not possible to generally determine the date of such transitions\footnote{Consequently they are excluded from the analysis alltogether, and thus also not contributing to the the hazard rate for closing, despite being entities at risk for closing. However, as their numbers are limited to 25 entries in the database, the bias this introduces into closing risks is likely not substantial.}
The main issues that made the determination of the date impossible was the intransparency of these organizations (\texttt{event in the past, no track of organizational history}), as well as the gradual process.
As the presence of non-founder members does not categorically exclude a museum from being categorized as private museum as long as third-party influence is limited, this influence can grow until the museum bears little resemblance to its former self.
The possibility of other outcomes than surival and closures has been investigated for NPOs more generally
\parencite{Searing_2020_zombies,HernandezOrtiz_2022_discontinuity,Helmig_Ingerfurth_Pinz_2013_nonprofit}, and for private museums specifically: 
\textcite{Walker_2019_collector} argues that in the case of Germany, private museum founders turn to the state to ensure the survival of their projects in the form of public-private partnerships.
Investigating this process of transformation into a different organizational form thus provides fruitful avenue for further research. 
\paragraph*{closing reasons not as rich as previous studies}

The extent to which closing factors are investigated does not reach the level of detail by \cite{Velthuis_Gera_2024_fragility}, who investigate each closing separately.
On the other hand, the inclusion of the museums who do not close results in a more more complete comparatison. 
A research approach that would combine the strengths of both approaches could for example consist of leveraging additional data sources, such as financial histories of founder companies to investigate financial issues or search engine results, media reports or social media posts to assess interest from the public.

founder age bracket

environment can be measured more
\begin{itemize}
\item geocoding of other institutiosn
\item tax incentives
\item government spending
\item government: political changes, censorship, persecution
\end{itemize}
-> this might futher specify the niches in which PMs can endure

remain intransparent

Museum characteristics might influence closing to a larger extent than has been considered here.
For example, museum size, facilities, as well as marketing and outreach activities might be related to long-term survival prospect.
For example, museums with more XXX might be more expensive to run, but on the other hand attract higher visitor numbers, cooperations and sponsorships. \texttt{liability of smallness}
Furthermore, having XXX might reflect founder commitment/resources.
While the original data collection included the measurement of various museum characteristics, such as floor and collection size, facilities, and cooperations, it became clear in the process that desk research was unsuitable to construct reliabile indicators from these findings due to difficulty in assessing them in many cases.
Especially for closed museums data collection proved difficult, as in many cases websites had been shut down, and while in many cases snapshots of the main website were available in the Internet archive, \texttt{subdomains} were archived much less frequently, and sometimes inaccessible despite being archived due to changes in webtechnology (such as flash).
While it would be possible to
absence of evidence as evidence of absence

\textbf{supplementary material}

covid effect may take longer to be effective
\subsubsection*{future research}

The US register data on NPO tax filings constitutes a promising source for data for museum research (while this data source has seen frequent usage in non-profit management, it remains underexploited in the field of museum research). 
While the United States private museum population with XX museums and YY closings might not be sufficient to test whether private museums have a a higher closing risks than other museums, it still hosts a large population of museums.
Thus this data can be used to establish average mortality estimates for museums generally as well as different types of museums, which can be compared to those observed for private museums globally. 
Its furthermore a rich source of financial data

egmus doesn't have that

>>>>

% Created 2024-09-11 wo 11:36
% Intended LaTeX compiler: pdflatex
\documentclass[12pt]{article}

@@

Neither gender of the founder nor external shocks (the Great Recession and the Covid-19 Pandemic) are associated with significantly higher closing rates.
Given the non-significant gender difference, it can be expected that the population of private museums will continue to remain a man-dominated affair, unless a strong increase in museum founding by women takes place.
At the same time, this finding also provides no evidence that female founders face gender-specific challenges, at least not to an extent that these decrease the survival prospects of their private museums.
The insignificance of the coefficients of the external shocks indicate that museums are overall relatively resilient; possibly unlike other types of art organizations with lower assets and higher reliance on earned incomes \parencite[p.102]{Bowen_etal_1994_charitable}.   
\paragraph*{Founder dependence (H2)}

@@

While so far little evidence points toward low diversification, i.e. reliance on the founder, as a key vulnerability of private museums to unvoluntary closures, the centrality of the founder may be more pronounced in voluntary closures. 
Indicated by unconventional naming choice, a subset of founders might from the very beginning envision the museum as temporary.
On the other hand, higher closing rates for unconventionally named museums might reflect absence of recognition and support by third parties in the environment.
A dimensions where the influence of the environment can however be ascertained with greater certainty is the role of audiences and other private museums.
While the overall relationship is complex, for the majority of private museums increases in both variables are associated with higher closing rates.
Whereas it is plausible that the association between higher institution counts and closing rates reflects competitive pressures, it is less obvious what may lay behind observing a positive association behind larger population and closing risk. 
\subsection*{Conclusion}

\subsubsection*{Limitations}

@@

While overall private museum closing rates are generally not diverging from museums in general, it has to be kept in mind that this comparison is based on American non-profit museums.
To my knowledge, currently no data sources exist that allow estimation of closing rates of (European) public, state-funded museums.
Although I argue that the American non-profit museums serve as a better comparison to evaluate private museum surival rates than state-funded museums because the comparison allows to focus on the centrality of the founder, the contrast to public museums would still provide additional insights into the relationship between ownership and survival rates.
While public museums have been characterized to some extent as very stable due to government picking up deficits \parencite{Meier_Frey_2003_faces,Bechtler_Imhof_2018_future}, \textcite{Walker_2019_collector} argues that due to declining cultural spending even public museums might face hard times.
Most national public flagship institutions might remain sufficiently funded to survive, however this may not be the case for the much larger of smaller museums in less touristic regions who rely on more limited municipal funding (49\% of the European museum population, \cite{EGMUS_2024_complete}), as \textcite{StylianouLambert_etal_2014_museums} explore for a number of smaller publicly funded ethnography museums in Cyprus.
Yet without access to museum registers comparable to the IRS/NCCS datasets, such questions of the relation between a wider range of organizational form and survival remain speculative.

The primary mechanism that has been argued to lead to museum closure is financials, and while this study measures it indirectly, primarily through indicators on the level of the founder, museum and environment, the financial situation of a museum is not used directly.
The primary reason for this is data availability.
While centralized non-profit financial information is available for American non-profits through the IRS/NCCS files, the much larger private museum population in the rest of the world is not subject to similar reporting requirements.
The effort to compile such information on a global scale ranges from extremely challenging (harmonizing different reporting standards, language barriers) to outright impossible for institutions who do not even (have to) report public financial information.
Similarly, other museum characteristics, such as as museum size, facilities, as well as marketing and outreach activities might be related to long-term survival prospect\footnote{While the original data collection included the assessment of various museum characteristics, such as floor and collection size, facilities, and cooperations, it became clear during data analysis that the data collected was unsuitable to construct accurate indicators from these findings due to difficulty in assessing them in many cases.
Especially for closed museums data collection proved difficult, as in many cases websites had been shut down, and while in many cases snapshots of the main website were available in the Wayback machine, less central pages (such as those with information about facilities and activities), were archived much less frequently, and sometimes inaccessible despite being archived due to changes in webtechnology (such as the retirement of Adobe flash).}
As such, private museums remain relatively intransparent, which does not prevent, but to some extent limits the identification of risk factors.

For this study, I only investigated the occurence of closure events.
However, private museums can stop existing as private museums not just by closing down, but also by being transformed into other organizational forms.
One way in which this can happen is when other parties, such as local governments other philanthropists, foundations, or corporations become substantially involved in the ownership, governance and/or funding of the museum.
Sometimes this involves changes in the composition of the board of directors (or equivalent), with new parties gaining seats and hence influence in decision-making, thereby diminishing the private character of the museum to an extent that it no longer fits the working definition.
Museums which experienced this trajectory have been excluded from the analysis as explorative investigation showed that it was not possible to consistently determine the date of such transitions\footnote{Consequently they are excluded from the analysis alltogether, and thus also not contributing to the the hazard rate for closing, despite being entities at risk for closing for at least some time. However, as their numbers are limited to 25 entries in the database, the bias this introduces into closing risks is likely not substantial.}
The main issues that made the determination of the date impossible was the intransparency of these organizations, the gradual process of the transition, as well as that these events life often years or decades in the past.
The possibility of other outcomes than surival and closures has been investigated for NPOs more generally
\parencite{Searing_2020_zombies,HernandezOrtiz_2022_discontinuity,Helmig_Ingerfurth_Pinz_2013_nonprofit}, and for private museums specifically on a case study basis: 
\textcite{Walker_2019_collector} argues that in the case of Germany, private museum founders turn to the state to ensure the survival of their projects in the form of public-private partnerships.
Investigating this process of transformation into a different organizational form thus provides fruitful avenue for further research.

While this study has already included measures of both the national and local environment, the ability to clearly observe spatial museum location allows to measure environmental influences in greater detail.
For once, so far competition has been limited to other private museums, but geocoding of a wider range of institutions would lead to a more encompassing assessment of competitive and dynamics.
Furthermore, including wider institutional changes such as tax incentives, government subsidies, but also political changes and censorship would lead to a more complete description of risk factors, and in more detail identify niches which are particularly suited for private museums.
\subsubsection*{future research}

The US register data on NPO tax filings constitutes a promising source for data for museum research (while this data source has seen frequent usage in non-profit management, it remains underexploited in the field of museum research). 
While the United States private museum population with 70 museums and 9 closings might not be sufficient to test whether private museums have a a higher closing risks than other museums, it still hosts a large population of museums.
Thus this data can be used to establish average mortality estimates for museums generally as well as different types of museums, which can be compared to those observed for private museums globally. 
The availability of financial data for American institutions through the NCCS/IRS may also provide tentative insights into the finances of private museums.
